Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/abstract.tex
checkpoint comments from frans through section 2

\textit{Nail} is an interface generator that greatly reduces the
programmer effort for safely parsing and generating protocols defined
by a Parser-Expression based grammar.  Nail introduces several key
ideas to achieve its goal.  First, Nail uses the protocol grammar to
define not just the data format, but also the internal object model
of the data.  Second, Nail eliminates the notion of semantic actions,
used by existing parser generators, which reduces the expressive power
but allows Nail to both {\em parse} data formats and {\em generate}
them from the internal object model, by maintaining a {\em semantic
bijection} between the data format and the object model.  Third,
Nail introduces {\em dependency fields} and {\em stream transforms} to
capture protocol features such as size and offset fields, checksums,
and compressed data, which are impractical to express in existing
protocol languages.  Using Nail, we implement a complete DNS server in
C in under 300 lines of code and grammar, and a complete {\tt unzip}
program in C in 220 lines of code and grammar, demonstrating that Nail
makes it easy to parse complex real-world data formats.

>>>>

\textit{Nail} is an interface generator that greatly reduces the
programmer effort for safely parsing and generating protocols defined
by a Parser-Expression based grammar.
\XXX[what is an interface generator?]
\XXX[what is a Parser-Expression based grammar?]
Nail introduces several key
ideas to achieve its goal.  First, Nail uses the protocol grammar to
define not just the data format, but also the internal object model
of the data.  Second, Nail eliminates the notion of semantic actions,
used by existing parser generators, which reduces the expressive power
but allows Nail to both {\em parse} data formats and {\em generate}
them from the internal object model, by maintaining a {\em semantic
bijection} between the data format and the object model.  Third,
Nail introduces {\em dependency fields} and {\em stream transforms} to
capture protocol features such as size and offset fields, checksums,
and compressed data, which are impractical to express in existing
protocol languages.  Using Nail, we implement a complete DNS server in
C in under 300 lines of code and grammar, and a complete {\tt unzip}
program in C in 220 lines of code and grammar, demonstrating that Nail
makes it easy to parse complex real-world data formats.  Performance
experiments show that a Nail-based DNS server can outperform the widely
used BIND DNS server, demonstrating that Nail parsers can achieve good
performance.

### 2014-05-01 paper/osdi14/intro.tex
add an explicit "motivation section"
current CVE vulnerability analysis from the security eval should move there.
checkpoint comments from frans through section 2

A promising approach to avoid such vulnerabilities is to specify
a precise grammar for the input data format, and to use a parser
generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
processing code.  Developers that use a parser generator do not need
to write error-prone input processing code on their own, and as long
as the parser generator is bug-free, the application will be safe from
input processing vulnerabilities.  Unfortunately, applying this approach
in practice, using state-of-the-art parser generators, still requires
too much manual programmer effort, making it error-prone, as we describe
next.

First, parser generators typically parse inputs into an abstract syntax
tree (AST) that corresponds to the grammar.  In order to produce a data
structure that the rest of the application code can easily process,
application developers must write explicit {\em semantic actions} that
update the application's internal representation of the data based on
each AST node.  Writing these semantic actions by hand is error-prone,
much like other input processing code, and mistakes can result in memory
corruption bugs or misinterpreted inputs.  Writing these semantic actions
also requires the programmer to describe the structure of the input three
times---once to describe the grammar, once to describe the internal data
structure, and once again in the semantic actions that translate the
grammar into the data structure---leading to another potential source
of bugs and inconsistencies.

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators focus on just parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is error-prone.  Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
into an AST\@.

@@

Finally, real-world data formats, for example PNG or PDF,  are hard to represent with
existing parser generators. Those parsers cannot directly deal with length or checksum fields
directly, so the programmer has to either write potentially unsafe code to deal with such features,
or build contrived grammar constructs, such as introducing one grammar rule for each possible value
of a length field. Offset fields usually require the programmer to manipulate a parser's internal
state to re-position its input. More complicated transformations, such as
handling compressed data, cannot be represented at all.

@@

Third, Nail introduces two abstractions, \emph{dependent fields} and \emph{transformations}, to
elegantly handle problematic structures, such as offset fields or checksums. Dependent fields are
fields in the protocol whose value depends in some way on the value or layout of other parts of the
format. Transformations allow the programmer to escape the generated code to modify the raw data and
interact with dependent fields in a controlled manner.

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for ZIP files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to parse and generate external
data representations.

The rest of this paper is organized as follows.  \S\ref{s:relwk}
puts Nail in the context of related work.  \S\ref{s:design} describes
Nail's design.  \S\ref{s:impl} discusses our initial implementation
of Nail.  \S\ref{s:eval} provides evaluation results.
\S\ref{s:future} suggests several directions for future work, and
\S\ref{s:concl} concludes.

>>>>

A promising approach to avoid such vulnerabilities is to specify
a precise grammar for the input data format, and to use a parser
generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
processing code.  Developers that use a parser generator do not need
to write error-prone input processing code on their own, and as long
as the parser generator is bug-free, the application will be safe from
input processing vulnerabilities.

Unfortunately, applying this approach in practice, using state-of-the-art
parser generators, still requires too much manual programmer effort,
and is still error-prone, for four reasons:

First, parser generators typically parse inputs into an abstract syntax
tree (AST) that corresponds to the grammar.  In order to produce a data
structure that the rest of the application code can easily process,
application developers must write explicit {\em semantic actions} that
update the application's internal representation of the data based on
each AST node.  Writing these semantic actions by hand is error-prone,
much like other input processing code, and mistakes can result in memory
corruption bugs or misinterpreted inputs.
\XXX[do we have any evidence to cite for this claim?]
Writing these semantic actions
also requires the programmer to describe the structure of the input three
times---once to describe the grammar, once to describe the internal data
structure, and once again in the semantic actions that translate the
grammar into the data structure---leading to another potential source
of bugs and inconsistencies.

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators focus on just parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is error-prone.
\XXX[do we have any evidence to cite for this claim?]
Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
into an AST\@.

@@

Finally, real-world data formats, for example PNG or PDF, are hard to
represent with existing parser generators.  Those parsers cannot directly
deal with length or checksum fields directly, so the programmer has to
either write potentially unsafe code to deal with such features, or build
contrived grammar constructs, such as introducing one grammar rule for
each possible value of a length field.  Offset fields, which specify
the position at which some data structure is located, usually require
the programmer to manipulate a parser's internal state to re-position
its input.  More complicated transformations, such as handling compressed
data, cannot be represented at all.

@@

Third, Nail introduces two abstractions, \emph{dependent fields} and
\emph{transformations}, to elegantly handle problematic structures,
such as offset fields or checksums.  Dependent fields capture fields
in a protocol whose value depends in some way on the value or layout
of other parts of the format; for example, offset or length fields,
which specify the position or length of another data structure, fall
in this category.  Transformations allow the programmer to escape the
generated code to modify the raw data and interact with dependent fields
in a controlled manner.

To evaluate whether Nail's design is effective at handling real-world
data formats, we implemented a prototype of Nail for C\@.  Using our
prototype, we implemented succinct grammars for parts of an IP network
stack, for DNS packets, and for ZIP files, each in about a hundred lines
of grammar.  On top of these grammars, we were able to build a fully
functional DNS server in under 200 lines of C code, and an {\tt unzip}
utility in about 50 lines of C code.  This suggests both that Nail is
effective at handling complex real-world data formats, and that Nail
makes it easy for application developers to parse and generate external
data representations.  Performance results show that the Nail-based DNS
server outperforms the widely used BIND DNS server, demonstrating that
Nail-based parsers and generators can achieve good performance.

The rest of this paper is organized as follows.  \S\ref{s:relwk}
puts Nail in the context of related work.
\S\ref{s:motivation} motivates the need for Nail by examining
past data format vulnerabilities.
\S\ref{s:design} describes
Nail's design.  \S\ref{s:impl} discusses our initial implementation
of Nail.  \S\ref{s:eval} provides evaluation results.
\S\ref{s:future} suggests several directions for future work, and
\S\ref{s:concl} concludes.

### 2014-05-01 paper/osdi14/relwk.tex
checkpoint comments from frans through section 2

Generating parsers and generators from an executable specification is the
core concept of interface generators, such as CORBA~\cite{omg:corba},
XDR~\cite{rfc:1832}, or Protocol Buffers~\cite{varda2008}.  However,
while interface generators work well for defining new protocols, they
do not provide full control over the data format, and therefore cannot
be used to implement existing protocols and data formats.

@@

Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
generating output for OCaml.  Parsifal structures grammars as an OCaml
type that holds an internal model and functions for parsing input and
output.  However, their generator can produce parsers and generators
only for simple, fixed-size structures.  The programmer can then use
these when implementing parsers and generators for more complicated
formats, manually handling offsets, checksums, and the like, risking bugs.
Nail handles more complicated constructs without the programmer manually
writing code to support them.

>>>>

Generating parsers and generators from an executable specification is the
core concept of interface generators, such as CORBA~\cite{omg:corba},
XDR~\cite{rfc:1832}, and Protocol Buffers~\cite{varda2008}.  However,
interface generators do not allow the programmer to specify the byte-level
data format; instead, they define their own data encoding that is
specific to a particular interface generator.  For instance, XDR-based
protocols are incompatible with Protocol Buffers.  Moreover, this means
that interface generators cannot be used to interact with existing protocols
that were not defined using that interface generator in the first place.
As a result, interface generators cannot be used to parse or generate
widely used formats such as DNS or ZIP, which is a goal for Nail.

@@

Parsifal~\cite{ANSSI:parsifal} is a parser framework that also supports
generating output for OCaml.  Parsifal structures grammars as an OCaml
type that holds an internal model and functions for parsing input and
output.  However, Parsifal can produce parsers and generators
only for simple, fixed-size structures.  The programmer can then use
these when implementing parsers and generators for more complicated
formats, manually handling offsets, checksums, and the like, risking bugs.
Nail handles more complicated constructs without the programmer manually
writing code to support them.
